---
title: "Final_project"
format: 
  revealjs:
   css: styles.css
editor: visual
number-sections: true
toc: true
---

# Intro

**text here** 

# Literature review

## Determining Factors in MOOCs Completion Rates: Application Test in Energy Sustainability Courses

**Introduction:**This study explores factors affecting Massive Open Online Course (MOOC) completion rates, specifically in energy sustainability courses. It examines how personal, family, social, labor, and instructional design factors influence engagement and course completion.

**Methods:** Data from 8,737 participants in 12 MOOCs (2017-2018) were analyzed. Surveys assessed participants’ motivations, backgrounds, and interactions to identify key factors influencing enrollment and completion.

**Results:** Personal, family, social, labor, and instructional design factors significantly impact MOOC engagement and completion. "Professional development" was most influential, particularly for technical graduates. Effective instructional design, including interactive elements and course relevance, was vital for younger participants, while traditional xMOOC design correlated with lower completion rates.

**Conclusions:** Improving MOOC completion rates hinges on effective instructional design, strong personal motivation, and professional relevance. These insights are essential for educators and course designers seeking to boost MOOC effectiveness and engagement.

## Massive Open Online Course Completion Rates Revisited: Assessment, Length and Attrition

**Introduction:** This study updates Jordan's 2014 work by analyzing enrollment and completion data from 221 MOOCs to identify factors affecting completion rates amidst changing enrollment trends.

**Methods:** The analysis uses multiple regression to evaluate how course length, start date, and assessment type impact completion rates, with data showing rates from 0.7% to 52.1%.

**Results:** Longer courses and those using peer grading are associated with lower completion rates. Course length and early engagement are significant factors, while university ranking and total enrollment are not.

**Conclusions:** Course length and assessment type significantly affect MOOC completion rates. Further research is needed due to data limitations and evolving trends.


**text here**

# Hypothesis

**text here**

# Methods
**In this project, we undertook a thorough analysis by first cleaning the data to ensure accuracy. We then performed correlation tests to identify relationships between variables, and checked for normal distribution using the Shapiro-Wilk test. Our analysis included running ANOVA to compare group means and using the Friedman test as a non-parametric alternative. We also calculated various means to summarize the data and created graphs to visualize the connections between different variables. This approach allowed us to gain a clear understanding of the relationships within the data.**

##lists of statistical tests
**1.

## Loading the necessary libraries

```{r, results='hide'}
library(dplyr)
library(tidyr)
library(tidyverse)
```

```{r}
# Reading the file into R
data = read.csv("data.csv")
```

```{r}
# Convert to tibble
data = as_tibble(data)
```

## Inspecting and cleaning the data

```{r}
# Insepecting the data
summary(data)
```

```{r}
#Check for NA's
sapply(data, function(x) sum(is.na(x)))
# NO NA's
```

```{r}
# Remove UserID column
data = select(data, -UserID)
```

```{r}
# Check for duplicates
duplicates = data[duplicated(data), ]
duplicates 
# We can see that there are 877 rows of duplicates
```

```{r}
#Removing duplicates
data = data |>
  distinct()
```

```{r}
#Check to see if the duplicates were removed, should equal to 0
sum(duplicated(data)) # 0 
```

```{r}
# count how many students from each category
category = data |>
  group_by(CourseCategory) |>
  summarise(Count = n())
category
```

## sampling 1500 from each category and normalzing the data

```{r}
set.seed(123)

data <- data %>%
  group_by(CourseCategory) %>%
  sample_n(size = 1500, replace = TRUE) %>%
  ungroup()


normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

data = data |>
  mutate(TimeSpentOnCourse = normalize(TimeSpentOnCourse),
         NumberOfVideosWatched = normalize(NumberOfVideosWatched),
         NumberOfQuizzesTaken = normalize(NumberOfQuizzesTaken),
         QuizScores = normalize(QuizScores),
         CompletionRate = normalize(CompletionRate))

data <- data %>%
  mutate(across(where(is.numeric) & !all_of(c("DeviceType")), round, 3))
data
```
##shapiro test 
**testing to see if the data is normally distributing**

##anova 

##fridman 

##correlation

## Mean by category

```{r, results='hide', echo=TRUE}
data_health = data|>
  filter(CourseCategory == 'Health')

data_arts = data |>
  filter(CourseCategory == 'Arts')

data_science <- data %>%
  filter(CourseCategory== 'Science')

data_Programming <- data %>%
  filter(CourseCategory == 'Programming')

data_business <- data %>%
  filter(CourseCategory == 'Business')


calculate_mean_completion <- function(data) {
  mean_completion <- sum(data$CourseCompletion == 1) / 1500
  return(mean_completion)
}

# Calculate mean completion rates for each category
mean_completion_health <- calculate_mean_completion(data_health)
mean_completion_science <- calculate_mean_completion(data_science)
mean_completion_Programming <- calculate_mean_completion(data_Programming)
mean_completion_arts <- calculate_mean_completion(data_arts)
mean_completion_business <- calculate_mean_completion(data_business)

# Create a table with the results
completion_rates <- tibble(
  course = c("Health", "Science", "Programming", "Arts", "Business"),
  completion_rate = c(mean_completion_health, mean_completion_science, mean_completion_Programming, mean_completion_arts, mean_completion_business)
)


completion_rates_per = completion_rates %>%
  mutate(completion_rate = completion_rate * 100)

# Create the bar chart
ggplot(completion_rates_per, aes(x = course, y = completion_rate, fill = course)) +
  geom_bar(stat = 'identity') +
  geom_text(aes(label = round(completion_rate, 1)), vjust = -0.5, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  labs(title = "Mean of Course Completion by Category",
       x = "Course Category",
       y = "completion of course (pass/not pass) (%)") +
  theme_minimal() +
  theme(legend.position = "none")
```
## mean by variables 

##mean by device type 

##plotting the data

#Results

## 

**code, pics, text here** \>\>\>\>\>\>\> 19609bc9b7f76d53fa40bfd18e1666f5365f5327

# Conclusions

**text here**

# later put in methods

# Loading the necessary libraries

```{r, results='hide', message=FALSE}
library(dplyr)
library(tidyr)
library(tidyverse)
```

```{r}
# Reading the file into R
data = read.csv("data.csv")
```

```{r}
# Convert to tibble
data = as_tibble(data)
```

```{r}
# Insepecting the data
summary(data)
```

```{r}
#Check for NA's
sapply(data, function(x) sum(is.na(x)))
# NO NA's
```

```{r}
# Remove UserID column
data = select(data, -UserID)
```

# Inspecting and cleaning the data

```{r}
# Check for duplicates
duplicates = data[duplicated(data), ]
duplicates 
# We can see that there are 877 rows of duplicates
```

```{r}
#Removing duplicates
data = data |>
  distinct()
```

```{r}
#Check to see if the duplicates were removed, should equal to 0
sum(duplicated(data)) # 0 
```

```{r}
# count how many students from each category
category = data |>
  group_by(CourseCategory) |>
  summarise(Count = n())
category
```

## sampling 1500 from each category and normalzing the data

```{r}
set.seed(123)

data <- data %>%
  group_by(CourseCategory) %>%
  sample_n(size = 1500, replace = TRUE) %>%
  ungroup()


normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

data = data |>
  mutate(TimeSpentOnCourse = normalize(TimeSpentOnCourse),
         NumberOfVideosWatched = normalize(NumberOfVideosWatched),
         NumberOfQuizzesTaken = normalize(NumberOfQuizzesTaken),
         QuizScores = normalize(QuizScores),
         CompletionRate = normalize(CompletionRate))

data <- data %>%
  mutate(across(where(is.numeric) & !all_of(c("DeviceType")), round, 3))
data
```

## Mean by category

```{r, eval=FALSE, include=FALSE}
data_health = data|>
  filter(CourseCategory == 'Health')

data_arts = data |>
  filter(CourseCategory == 'Arts')

data_science <- data %>%
  filter(CourseCategory== 'Science')

data_Programming <- data %>%
  filter(CourseCategory == 'Programming')

data_business <- data %>%
  filter(CourseCategory == 'Business')


calculate_mean_completion <- function(data) {
  mean_completion <- sum(data$CourseCompletion == 1) / 1500
  return(mean_completion)
}

# Calculate mean completion rates for each category
mean_completion_health <- calculate_mean_completion(data_health)
mean_completion_science <- calculate_mean_completion(data_science)
mean_completion_Programming <- calculate_mean_completion(data_Programming)
mean_completion_arts <- calculate_mean_completion(data_arts)
mean_completion_business <- calculate_mean_completion(data_business)

# Create a table with the results
completion_rates <- tibble(
  course = c("Health", "Science", "Programming", "Arts", "Business"),
  completion_rate = c(mean_completion_health, mean_completion_science, mean_completion_Programming, mean_completion_arts, mean_completion_business)
)


completion_rates_per = completion_rates %>%
  mutate(completion_rate = completion_rate * 100)

# Create the bar chart
ggplot(completion_rates_per, aes(x = course, y = completion_rate, fill = course)) +
  geom_bar(stat = 'identity') +
  geom_text(aes(label = round(completion_rate, 1)), vjust = -0.5, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  labs(title = "Mean of Course Completion by Category",
       x = "Course Category",
       y = "Completion Rate (%)") +
  theme_minimal() +
  theme(legend.position = "none")
```

